---
title: "Data Analysis Report"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
# Set chunk options
knitr::opts_chunk$set(
  echo = TRUE,           # Show code by default
  warning = FALSE,       # Hide warnings
  message = FALSE,       # Hide messages
  fig.width = 10,        # Default figure width
  fig.height = 6,        # Default figure height
  fig.align = "center",  # Center figures
  cache = FALSE          # Don't cache results
)

# Load required packages
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

# Source utility functions
# source(here("scripts/utils.R"))
```

# Introduction

Brief overview of the analysis objectives and data sources.

## Objectives

- Objective 1
- Objective 2
- Objective 3

## Data Description

Describe your data here.

# Data Loading

```{r load-data}
# Load your data
# data <- read_csv(here("data/processed/cleaned_data.csv"))

# Display first few rows
# head(data) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# Exploratory Data Analysis

## Data Summary

```{r data-summary}
# Get summary statistics
# summary(data)
# skimr::skim(data)
```

## Visualizations

```{r visualizations}
# Create visualizations
# ggplot(data, aes(x = variable)) +
#   geom_histogram(bins = 30, fill = "steelblue") +
#   theme_minimal() +
#   labs(title = "Distribution of Variable")
```

# Statistical Analysis

## Descriptive Statistics

```{r descriptive-stats}
# Calculate descriptive statistics
# summary_table <- data %>%
#   group_by(group_var) %>%
#   summarise(
#     n = n(),
#     mean = mean(value_var, na.rm = TRUE),
#     sd = sd(value_var, na.rm = TRUE)
#   )
# 
# summary_table %>%
#   kable(digits = 2) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))
```

## Inferential Statistics

```{r inferential-stats}
# Perform statistical tests
# model <- lm(outcome ~ predictor, data = data)
# summary(model)
```

# Results

Present your key findings here.

# Discussion

Interpret your results and discuss implications.

# Conclusions

Summarize main conclusions and potential next steps.

# Session Information

```{r session-info}
sessionInfo()
```
