---
title: "Exploratory Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)

library(tidyverse)
library(skimr)
library(here)
```

# Overview

This notebook provides a comprehensive exploratory data analysis (EDA) workflow.

# Load Data

```{r load-data}
# Load your dataset
# data <- read_csv(here("data/raw/your_file.csv"))
```

# Data Structure

```{r structure}
# Check data structure
# glimpse(data)
# str(data)
```

# Data Quality

## Missing Values

```{r missing-values}
# Check for missing values
# missing_summary <- data %>%
#   summarise(across(everything(), ~sum(is.na(.)))) %>%
#   pivot_longer(everything(), names_to = "variable", values_to = "missing_count") %>%
#   mutate(missing_pct = round(missing_count / nrow(data) * 100, 2)) %>%
#   arrange(desc(missing_count))
# 
# missing_summary

# Visualize missing data
# missing_summary %>%
#   filter(missing_count > 0) %>%
#   ggplot(aes(x = reorder(variable, missing_pct), y = missing_pct)) +
#   geom_col(fill = "steelblue") +
#   coord_flip() +
#   labs(title = "Missing Data by Variable",
#        x = "Variable",
#        y = "% Missing")
```

## Duplicates

```{r duplicates}
# Check for duplicate rows
# n_duplicates <- sum(duplicated(data))
# cat("Number of duplicate rows:", n_duplicates)
```

# Univariate Analysis

## Numeric Variables

```{r numeric-summary}
# Summary statistics for numeric variables
# skim(data)
```

```{r numeric-distributions}
# Visualize distributions of numeric variables
# data %>%
#   select(where(is.numeric)) %>%
#   pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
#   ggplot(aes(x = value)) +
#   geom_histogram(bins = 30, fill = "steelblue") +
#   facet_wrap(~variable, scales = "free") +
#   theme_minimal() +
#   labs(title = "Distribution of Numeric Variables")
```

## Categorical Variables

```{r categorical-summary}
# Frequency tables for categorical variables
# data %>%
#   select(where(is.character)) %>%
#   map(~table(.))
```

```{r categorical-plots}
# Visualize categorical variables
# data %>%
#   select(where(is.character)) %>%
#   pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
#   count(variable, value) %>%
#   ggplot(aes(x = reorder(value, n), y = n)) +
#   geom_col(fill = "steelblue") +
#   facet_wrap(~variable, scales = "free") +
#   coord_flip() +
#   theme_minimal() +
#   labs(title = "Frequency of Categorical Variables",
#        x = "Category",
#        y = "Count")
```

# Bivariate Analysis

## Correlations

```{r correlations}
# Correlation matrix for numeric variables
# cor_matrix <- data %>%
#   select(where(is.numeric)) %>%
#   cor(use = "complete.obs")
# 
# cor_matrix

# Visualize correlation matrix
# library(corrplot)
# corrplot(cor_matrix, method = "color", type = "upper", 
#          tl.col = "black", tl.srt = 45)
```

## Relationships

```{r scatter-plots}
# Scatter plots for key relationships
# ggplot(data, aes(x = var1, y = var2)) +
#   geom_point(alpha = 0.5, color = "steelblue") +
#   geom_smooth(method = "lm", color = "red") +
#   theme_minimal() +
#   labs(title = "Relationship between Variables")
```

# Outlier Detection

```{r outliers}
# Identify outliers using box plots
# data %>%
#   select(where(is.numeric)) %>%
#   pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
#   ggplot(aes(x = variable, y = value)) +
#   geom_boxplot(fill = "steelblue") +
#   coord_flip() +
#   theme_minimal() +
#   labs(title = "Outlier Detection")
```

# Key Findings

Summarize the main insights from your EDA:

1. Finding 1
2. Finding 2
3. Finding 3

# Next Steps

Based on the EDA, outline the next steps for analysis.
